---
title: "p8105_hw1_sal2222"
author: "Stephen Lewandowski"
date: "2018-09-13"
output: 
  github_document:
    toc: true
---

```{r setup, include = FALSE}
library(tidyverse)
```

# Problem 1

This problem focuses on variable types and coercion.

First, I will create a data frame comprised of:
a random sample of size 10 from a uniform[0, 5] distribution,
a logical vector indicating whether elements of the sample are greater than 2,
a (length-10) character vector, and
a (length-10) factor vector.

```{r df_create}
set.seed(1234)

random_sample <- runif(10, min = 0, max = 5)

logical_GT2 <- random_sample > 2

character10 <- c("Data", "science", "is", "the", "use", "of", "data", "to",                       "formulate", "and")

factor10 <- factor(c("green", "red", "red", "blue", "green", "blue", "green",                         "red",   "red", "blue"))

problem1_df = tibble(
  random_sample,
  logical_GT2,
  character10, 
  factor10
 )

problem1_df

```

I will try to take the mean of each variable in my dataframe. 

```{r take_means}

mean(random_sample)

mean(logical_GT2)

mean(character10)

mean(factor10)

```

What works and what doesn’t? Why?

The random uniform distribution sample vector works because you can calculate the mean from the numeric class. The logical vector works because mean returns the proportion of "TRUE" observations. The charcter and factor vectors do not work. Both variables return "NA" because their class is neither numeric nor logical.


# Problem 2
