---
title: "p8105_hw1_sal2222"
author: "Stephen Lewandowski"
date: "2018-09-13"
output: 
  github_document:
    toc: true
---

```{r setup, include = FALSE}
library(tidyverse)
```

# Problem 1

This problem focuses on variable types and coercion.

## Create a dataframe

First, I will create a data frame with a random sample from a uniform distribution, a logical vector indicating whether elements of the sample are greater than 2, a character vector, and a factor vector.

```{r df_create}
set.seed(1234)

random_sample <- runif(10, min = 0, max = 5)

logical_GT2 <- random_sample > 2

character10 <- c("Data", "science", "is", "the", "use", "of", "data", "to", "formulate", "and")

factor10 <- factor(c("green", "red", "red", "blue", "green", "blue", "green", "red", "red", "blue"))

constructed_df = tibble(
  random_sample,
  logical_GT2,
  character10, 
  factor10
 )

constructed_df

```

## Variable means
I will attempt to take the mean of each variable in my dataframe. 

```{r take_means}

mean(random_sample)
mean(logical_GT2)
mean(character10)
mean(factor10)

```

*What works and what doesn’t? Why?*

The random uniform distribution sample vector works because you can calculate the mean from the numeric class. The logical vector works because mean returns the proportion of "TRUE" observations. The charcter and factor vectors do not work. Both variables return "NA" because their class is neither numeric nor logical.

## Apply as.numeric function
*Write a code chunk that applies the as.numeric function to the logical, character, and factor variables. What happens?*

```{r reclass, results = 'hide', warning = FALSE}

as.numeric(logical_GT2)
as.numeric(character10)
as.numeric(factor10)

```

For the logial variable, the `as.numeric` function converts "TRUEs" and "FALSEs" to "1s" and "0s", respectively.

The `as.numeric` function does not work for the character variable. "NAs" are introduced by coersion for each observation.

For the factor variable, the `as.numeric` function assigns an integer for each unique level.  In my example, "blue" is changed to "1", "green" is changed to "2", and "red" is changed to "3".


## Multiple class conversions
*Convert your character variable from character to factor to numeric; similarly, convert your factor variable from factor to character to numeric. What happens?*

```{r reclass_multiple, results = 'hide', warning = FALSE}

as.numeric(as.factor(character10))
as.numeric(as.character(factor10))

```

Converting my character variable from character to factor to numeric assigned a unique level to each character and converted the levels to integers from 1 to 10.

Converting my factor variable from factor to character to numeric returns "NAs" by coersion for each observation since the character strings cannot be converted to numeric values. 

```{r {r reclass_multiple_output, echo = FALSE, warning = FALSE}

as.numeric(as.factor(character10))
as.numeric(as.character(factor10))

```


# Problem 2
