---
title: "p8105_hw1_sal2222"
author: "Stephen Lewandowski"
date: "2018-09-13"
output: 
  github_document:
    toc: true
---

```{r setup, include = FALSE}
library(tidyverse)
```

# Problem 1

This problem focuses on variable types and coercion.

## Create a mixed class dataframe

First, I will create a data frame with a random sample from a uniform distribution, a logical vector indicating whether elements of the sample are greater than 2, a character vector, and a factor vector.

```{r df_create}
set.seed(1234)

random_sample <- runif(10, min = 0, max = 5)

logical_GT2 <- random_sample > 2

character10 <- c("Data", "science", "is", "the", "use", "of", "data", "to", "formulate", "and")

factor10 <- factor(c("green", "red", "red", "blue", "green", "blue", "green", "red", "red", "blue"))

constructed_df <- tibble(
  random_sample,
  logical_GT2,
  character10, 
  factor10
 )

constructed_df

```

## Variable means
I will attempt to take the mean of each variable in my dataframe. 

```{r take_means}

mean(random_sample)
mean(logical_GT2)
mean(character10)
mean(factor10)

```

*What works and what doesn’t? Why?*

The random uniform distribution sample vector works because you can calculate the mean from the numeric class. The logical vector works because mean returns the proportion of "TRUE" observations. The charcter and factor vectors do not work. Both variables return "NA" because their class is neither numeric nor logical.

## Apply `as.numeric` function
*Write a code chunk that applies the as.numeric function to the logical, character, and factor variables. What happens?*

```{r reclass, results = 'hide', warning = FALSE}

as.numeric(logical_GT2)
as.numeric(character10)
as.numeric(factor10)

```

For the logical variable, the `as.numeric` function converts "TRUEs" and "FALSEs" to "1s" and "0s", respectively.

The `as.numeric` function does not work for the character variable. "NAs" are introduced by coersion for each observation.

For the factor variable, the `as.numeric` function assigns an integer for each unique level.  In my example, "blue" is changed to "1", "green" is changed to "2", and "red" is changed to "3".


## Multiple class conversions
*Convert your character variable from character to factor to numeric; similarly, convert your factor variable from factor to character to numeric. What happens?*

```{r reclass_multiple, results = 'hide', warning = FALSE}

as.numeric(as.factor(character10))
as.numeric(as.character(factor10))

```

Converting my character variable from character to factor to numeric assigned a unique level to each character and converted the levels to integers from 1 to 10.

```{r {r reclass_multiple_output_character, echo = FALSE, warning = FALSE}

as.numeric(as.factor(character10))

```

Converting my factor variable from factor to character to numeric returns "NAs" by coersion for each observation since the character strings cannot be converted to numeric values. 

```{r {r reclass_multiple_output_factor, echo = FALSE, warning = FALSE}

as.numeric(as.factor(character10))

```

# Problem 2

This problem focuses on plotting and the use of inline R code.

## Create a mixed class data frame

I will create a data frame comprised of two random samples from a standard Normal distribution, a logical vector indicating whether the sum of the random samples is greater than 0, and numeric and factor vectors derived from the logical vector. The first 10 rows of the data frame are displayed.

```{r create_5_var_df}
set.seed(1)
x <- rnorm(1000)

set.seed(2)
y <- rnorm(1000)

logical_sum_xy <- x + y > 0

numeric_from_logical <- as.numeric(logical_sum_xy)

factor_from_logical <- as.factor(logical_sum_xy)

five_variable_df <- tibble(
  x,
  y,
  logical_sum_xy,
  numeric_from_logical,
  factor_from_logical
)

five_variable_df

```

## Description of dataset 

*The following is a short description of my data, including:  the size of the dataset, the mean and median of x, the proportion of cases for which the logical vector is TRUE.*

The data frame `five_variable_df` has `r nrow(five_variable_df)` rows and `r ncol(five_variable_df)` columns.

The length of variable `x` is `r length(five_variable_df$x)`. The mean of `x` is `r mean(five_variable_df$x)`. The median of `x` is `r median(five_variable_df$x)`.    

The proportion of cases for which the logical vector `logical_sum_xy` is "TRUE" is `r mean(five_variable_df$logical_sum_xy)`. 



## Scatterplots of y vs x



Next I will show a scatterplots of `y` vs `x`, with points colored according to `logical_sum_xy`, `numeric_from_logical`, and `factor_from_logical`. 

### Colored by logical variable
```{r yx_scatter_logical}

plot_by_logical <- ggplot(five_variable_df, aes(x = x, y = y, color = logical_sum_xy)) + geom_point()
plot_by_logical
```

The color scale in the above plot displays teal for "TRUE" and orange for "FALSE", based on the logical classification of `x + y > 0`. 

### Colored by numeric variable

```{r yx_scatter_numeric}

plot_by_numeric <- ggplot(five_variable_df, aes(x = x, y = y, color = numeric_from_logical)) + geom_point()
plot_by_numeric

```

The color scale in the above plot displays light blue for values of "1" and dark blue for values of "0". Because the variable class is classified as "numeric", a continuous, graduated color scale is provided. 

### Colored by factor variable

```{r yx_scatter_factor}

plot_by_factor <- ggplot(five_variable_df, aes(x = x, y = y, color = factor_from_logical)) + geom_point()
plot_by_factor

```

The color scale in the above plot displays teal for "TRUE" and orange for "FALSE". The variable class is classified as "factor",  and the color scale uses two colors to reflect the two factor levels contained in the variable.



## Export scatterplot to project directory

I will now export my first scatterplot (colored by logical variable) to my project directory, with filename `scatterplot_yx_colorby_logical.png`.

```{r export_scatterplot}

ggsave(filename = "scatterplot_yx_colorby_logical.png", plot = plot_by_logical)

```




